Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-31 cav2017/cav2017.tex
nudge
extract lemmas from section 4.3

An overview of the controller synthesis is given in
Fig.~\ref{fig:CEGIS-precision-increment}.  One important observation is
that, for this approach, we verify and synthesize a controller over $k$ time
steps.  We~then compute a completeness threshold, $\overline{k}$, for this
controller, and verify correctness for $\overline{k}$ time steps. 
Essentially, $\overline{k}$ is the number of iterations required to
sufficiently unwind the closed-loop state-space system such that the
boundaries are not violated for any $k{>}\overline{k}$ (i.e., it is
sufficient to unwind the closed-loop state-space system up to $\overline{k}$
in order to ensure that $\phi_\mathit{safety}$ holds).

@@

 %\todo{unclear next: it seems like to select a single initial state here, rather than considering the set thereof - clarify. also, clarify relationship btw k and soundness. }

@@

This is relatively computationally expensive if the bounds on the
allowed initial states are large.  However, we can show that, if the
controller is safe for each of the corner cases of our hypercube of
allowed initial states, it is safe for any initial state in the
hypercube. Thus we only need to check $2^n$ initial states, where $n$
is the dimension of the state space (number of continuous variables).

To argue for this sufficient check, 
consider the set of initial states, $X_0$, which we assume to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. Then if $x_0=x$, we obtain 
\begin{align*}
x_k   &= (A_d - B_d K)^k x = (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i 
      = \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i = \sum_{i=1}^{2^n} \alpha_i x_k^i, 
 \end{align*}     
%
where $x_k^i$ denotes the trajectories obtained from the single vertex
$v_i$.  We~conclude that any $k$-step trajectory is encompassed, within a
convex set, by those generated from the vertices.

@@

\item The third {\sc verify} stage, {\sc complete}, checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate $K$ and
check that $k{\leq}\overline{k}$.  This is done according to the following
argument, illustrated in Fig.~\ref{fig:ct}.

A stable control system is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this in general not the case when dealing with complex
eigenvalues.  Consider the closed-loop matrix, which updates the states
every discrete time step, and select the eigenvalue $\vartheta$ with
smallest (non-trivial) imaginary value.  At every consecutive time steps
$kT_s, (k+1)T_s$, the dynamics projected on the corresponding eigenspace
rotate of $\vartheta T_s$ radiants.  Thus, taking $k{\leq}\overline{k}$ as the
ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k{\leq}\overline{k}$ steps
we have completed a full rotation, which results in a point closer to the
origin, and it is the completeness threshold.

@@

>>>>

An overview of the controller synthesis is given in
Fig.~\ref{fig:CEGIS-precision-increment}.  One important observation is
that, for this approach, we verify and synthesize a controller over $k$ time
steps.  We~then compute a completeness threshold, $\overline{k}$, for this
controller, and verify correctness for $\overline{k}$ time steps. 
Essentially, $\overline{k}$ is the number of iterations required to
sufficiently unwind the closed-loop state-space system such that the
boundaries are not violated for any $k{>}\overline{k}$.

\begin{lemma} There exists a $\overline{k}$ such that it is
sufficient to unwind the closed-loop state-space system up to $\overline{k}$
in order to ensure that $\phi_\mathit{safety}$ holds. 
\end{lemma}

\begin{proof}
A stable control system is known to have converging dynamics.  Assume the
closed-loop matrix eigenvalues are not repeated (which is sensible to do,
since we select them).  The distance of the trajectory from the reference
point (origin) decreases over time within subspaces related to real-valued
eigenvalues; however, this in general not the case when dealing with complex
eigenvalues.  Consider the closed-loop matrix, which updates the states
every discrete time step, and select the eigenvalue $\vartheta$ with
smallest (non-trivial) imaginary value.  At every consecutive time steps
$kT_s, (k+1)T_s$, the dynamics projected on the corresponding eigenspace
rotate of $\vartheta T_s$ radiants.  Thus, taking $k{\leq}\overline{k}$ as the
ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k{\leq}\overline{k}$ steps
we have completed a full rotation, which results in a point closer to the
origin, and it is the completeness threshold.
\end{proof}

@@

@@

@@

\item The third {\sc verify} stage, {\sc complete}, checks that the current
$k$ is large enough to ensure safety for any $k'{>}k$.  Here, we compute the
completeness threshold $\overline{k}$ for the current candidate $K$ and
check that $k{\leq}\overline{k}$.  This is done according to the argument given above
and illustrated in Fig.~\ref{fig:ct}.

@@

Checking that the safety specification holds for any initial state can be 
computationally expensive if the bounds on the
allowed initial states are large.

\begin{lemma} 
If a controller is safe for each of the corner cases of our hypercube of
allowed initial states, it is safe for any initial state in the
hypercube. 
\end{lemma}

Thus we only need to check $2^n$ initial states, where $n$
is the dimension of the state space (number of continuous variables). 
\begin{proof}
Consider the set of initial states, $X_0$, which we assume to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. Then if $x_0=x$, we obtain 
\begin{align*}
x_k   &= (A_d - B_d K)^k x = (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i 
      = \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i = \sum_{i=1}^{2^n} \alpha_i x_k^i, 
 \end{align*}     
%
where $x_k^i$ denotes the trajectories obtained from the single vertex
$v_i$.  We~conclude that any $k$-step trajectory is encompassed, within a
convex set, by those generated from the vertices.
\end{proof}
